---
title: "Introduction to R"
author: "Tan Sein Jone"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
---

1. Create a histogram for total pollution levels in 1990 with provinces in the x axis and total pollution levels in the y axis.

```{r}
library(ggplot2)

pollution <- read.csv("../../../python/python_2024/data/pollution.csv")
temperature <- read.csv("../../../python/python_2024/data/temperature.csv")

# Filter for the year 1990
df_1990 <- subset(pollution, Year == 1990)

# Drop the 'Year' column
df_1990 <- df_1990[, !(names(df_1990) %in% c("Year"))]

# Sum all pollution values for each province
df_1990$Total_Pollution <- rowSums(df_1990[, -1])

# Create a histogram with total pollution
ggplot(df_1990, aes(x = Province, y = Total_Pollution)) +
    geom_bar(stat = "identity") +
    ylab("Total Pollution") +
    ggtitle("Total Pollution in 1990") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

2. Create a histogram for total pollution and temperature levels in 1990 with provinces in the x axis and total pollution levels and temperature in the y axis. The x axis should contain full provincial names.

```{r}
library(dplyr)
library(tidyr)

names <- read.csv("../../../python/python_2024/data/province_names.csv")

# Map province names to pollution data
df_names <- names %>%
    right_join(pollution, by = c("Shortnam" = "Province"))

df_names <- df_names %>%
    rename(Province = Shortnam)

# Sum all pollution values
df_names <- df_names %>%
    mutate(Total_Pollution = rowSums(select(., 4:ncol(.))))

# Merge with temperature data
df_names <- df_names %>%
    inner_join(temperature, by = c("Province", "Year")) %>%
    select(-Province)

# Filter data for 1990
df_h <- df_names %>%
    filter(Year == 1990) %>%
    select(-Year)

# Determine the scaling factor
scaling_factor <- max(df_h$Total_Pollution) / max(df_h$temperature)

# Plot with ggplot2
ggplot(df_h, aes(x = Region)) +
    geom_bar(aes(y = Total_Pollution), stat = "identity", fill = "blue", width = 0.4, position = position_dodge(width = 0.4)) +
    geom_bar(aes(y = temperature * scaling_factor), stat = "identity", fill = "red", width = 0.4, position = position_dodge(width = 0.4)) +
    scale_y_continuous(
        name = "Total Pollution",
        sec.axis = sec_axis(~ . / scaling_factor, name = "Temperature")
    ) +
    labs(x = "Region", title = "Total Pollution and Temperature in 1990") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

3. Create a scatter plot for total pollution levels and temperature levels in 1990 with total pollution levels in the x axis and temperature levels in the y axis. Each point should represent a province.

```{r}
# Copy the dataframe
df_time <- pollution

# Calculate Total Pollution
df_time <- df_time %>%
    mutate(Total_Pollution = rowSums(select(., 3:ncol(.))))

# Set Province as the index
df_time <- df_time %>%
    select(Province, Year, Total_Pollution)

# Identify the province with the highest and lowest pollution in 1990
temp <- df_time %>%
    filter(Year == 1990)

max_province <- temp %>%
    filter(Total_Pollution == max(Total_Pollution)) %>%
    pull(Province)

min_province <- temp %>%
    filter(Total_Pollution == min(Total_Pollution)) %>%
    pull(Province)

filter_provinces <- c(max_province, min_province)

# Filter data for the provinces with the highest and lowest pollution
df_time <- df_time %>%
    filter(Province %in% filter_provinces)

# Calculate the change in pollution
df_time <- df_time %>%
    group_by(Province) %>%
    arrange(Year) %>%
    mutate(change = (Total_Pollution / lag(Total_Pollution)) - 1) %>%
    ungroup() %>%
    filter(!is.na(change))

# Prepare data for plotting
max_df <- df_time %>%
    filter(Province == max_province)

min_df <- df_time %>%
    filter(Province == min_province)

# Plot the change in pollution for the provinces with the highest and lowest pollution
ggplot() +
    geom_line(data = max_df, aes(x = Year, y = change, color = max_province), size = 1) +
    geom_line(data = min_df, aes(x = Year, y = change, color = min_province), size = 1) +
    labs(x = "Year", y = "Change in Total Pollution", title = "Change in Total Pollution Over Time") +
    scale_color_manual(values = c("blue", "red"), labels = c(max_province, min_province)) +
    theme_minimal() +
    theme(legend.title = element_blank())
```

4. Create a table with the average pollution levels for each industry by province. Provinces should be in rows and industries in columns. Province names should be full names. The table should also include the average temperature over time. Identify the industry with the highest average pollution level and the province with the highest temperature.

```{r}
# Calculate the mean pollution and temperature for each province
df_pivot <- pollution %>%
    group_by(Province) %>%
    summarise(across(3:8, mean, na.rm = TRUE))

# Calculate the sum of every column
column_sums <- df_pivot %>%
    select(-Province) %>%
    colSums()

df_temp <- temperature %>%
    group_by(Province) %>%
    summarise(Avg_Temperature = mean(temperature, na.rm = TRUE)) %>%
    left_join(names, by = c("Province" = "Shortnam"))

print(names(column_sums)[which.max(column_sums)])
print(df_temp$Region[which.max(df_temp$Avg_Temperature)])
```

5. Do a linear regression of total pollution levels on temperature levels. Plot the regression line on a scatter plot of total pollution levels and temperature levels.

```{r}
library(broom)

# Calculate the total pollution for each row
df_reg <- pollution %>%
    mutate(Total_Pollution = rowSums(select(., -Province)))

# Merge the pollution and temperature data frames
df_reg <- df_reg %>%
    inner_join(temperature, by = "Province") %>%
    select(Total_Pollution, temperature)

# Fit a linear regression model
model <- lm(Total_Pollution ~ temperature, data = df_reg)
summary(model)

# Plot the regression line
ggplot(df_reg, aes(x = temperature, y = Total_Pollution)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    labs(x = "Temperature", y = "Total Pollution") +
    theme_minimal()
```