---
title: "Introduction to R"
author: "Tan Sein Jone"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
---

# R and Academic Research

R is a powerful programming language and software environment for statistical computing and graphics. It is widely used in academic research for data analysis, statistical modeling, and visualization. In this tutorial, we will cover some basic concepts and techniques in R that are commonly used in academic research.

## Basic Regressions

One of the most common tasks in academic research is to estimate regression models. R provides a wide range of functions and packages for fitting different types of regression models. Here are some examples of basic regression models in R:

### Simple Linear Regression

Simple linear regression is used to model the relationship between a single independent variable and a continuous dependent variable. Here's an example of fitting a simple linear regression model in R:

#### Task 1: Fit a Simple Linear Regression Model to The mtcars Dataset

1. Load the `mtcars` dataset.
2. Fit a simple linear regression model with `mpg` as the dependent variable and `wt` as the independent variable.
3. Display the summary of the regression model.

```{r}
# Load the mtcars dataset
data(mtcars)

# Fit a simple linear regression model
model <- lm(mpg ~ wt, data = mtcars)

# Display the summary of the regression model
summary(model)
```

### Multiple Linear Regression

Multiple linear regression is used to model the relationship between multiple independent variables and a continuous dependent variable. Here's an example of fitting a multiple linear regression model in R:

#### Task 2: Fit a Multiple Linear Regression Model to The mtcars Dataset

1. Fit a multiple linear regression model with `mpg` as the dependent variable and `wt`, `hp`, and `qsec` as the independent variables.
2. Display the summary of the regression model.

```{r}
# Fit a multiple linear regression model
model <- lm(mpg ~ wt + hp + qsec, data = mtcars)

# Display the summary of the regression model
summary(model)
```

### Visualizing Regression Results

In addition to fitting regression models, it is also important to visualize the results to better understand the relationships between variables. Here's an example of visualizing the results of a simple linear regression model in R:

#### Task 3: Visualize the Results of The Simple Linear Regression Model

1. Plot the data points and the regression line.
2. Add the regression equation and the coefficient of determination (R-squared) to the plot.

```{r}
# Plot the data points and the regression line
plot(mtcars$wt, mtcars$mpg, xlab = "Weight", ylab = "Miles per Gallon", main = "Simple Linear Regression")
abline(model, col = "red")

# Add the regression equation and R-squared to the plot
text(5, 30, paste("Regression Equation: mpg = ", round(coef(model)[1], 2), "+", round(coef(model)[2], 2), "* wt"))
text(5, 28, paste("R-squared = ", round(summary(model)$r.squared, 2)))
```