---
title: "Light Data Work"
author: "Tan Sein Jone"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
---

# Built in Datasets

r comes with a number of built-in datasets that you can use to practice your data manipulation skills. You can view the list of datasets by running the following command:

```{r}
data()
```

You can access the datasets by running the following command:

```{r}
head(mtcars)
```

## Data Manipulation

### Task 1

1. Load the `mtcars` dataset.
2. Calculate the mean of the `mpg` column.
3. Sort the dataset by the `mpg` column in descending order.

```{r}
# Load the mtcars dataset
data(mtcars)

# Calculate the mean of the mpg column
mean(mtcars$mpg)

# Sort the dataset by the mpg column in descending order
mtcars_sorted <- mtcars[order(-mtcars$mpg), ]
head(mtcars_sorted)
```

### Task 2

1. For cars that have a `mpg` greater than theteh average, calculate the mean of the `hp` column.
2. Count the number of cars that have a `wt` greater than 2.

```{r}
# Calculate the mean of the hp column for cars with mpg greater than the average
mean(mtcars$hp[mtcars$mpg > mean(mtcars$mpg)])

# Count the number of cars with wt greater than 2
sum(mtcars$wt > 2)
```

### Task 3

1. Create a new column called `disp_per_cyl` that calculates the displacement per cylinder.
2. Calculate the mean of the `disp_per_cyl` grouped by the number of cylinders.

```{r}
# Create a new column called disp_per_cyl
mtcars$disp_per_cyl <- mtcars$disp / mtcars$cyl

# Calculate the mean of disp_per_cyl grouped by the number of cylinders
aggregate(mtcars$disp_per_cyl, by = list(mtcars$cyl), FUN = mean)
```

### Task 4

1. Create a new dataset that contains the average attribute for cars of every manufacturer.

```{r}
df <- mtcars
df$manufacturer <- sapply(strsplit(rownames(df), " "), "[", 1)
aggregate(df[, -1], by = list(df$manufacturer), FUN = mean)
```

## Data Visualization

### Task 5

1. Create a scatter plot of `mpg` vs `hp`.
2. Create a boxplot of `mpg` grouped by the number of cylinders.

```{r}
# Create a scatter plot of mpg vs hp
plot(mtcars$hp, mtcars$mpg, xlab = "Horsepower", ylab = "Miles per Gallon", main = "Miles per Gallon vs Horsepower")

# Create a boxplot of mpg grouped by the number of cylinders
boxplot(mpg ~ cyl, data = mtcars, xlab = "Number of Cylinders", ylab = "Miles per Gallon", main = "Miles per Gallon by Number of Cylinders")
```

### Task 6

1. Use the `ggplot2` package to create a scatter plot of `mpg` vs `hp`.

```{r}
# Load the ggplot2 package
library(ggplot2)

# Create a scatter plot of mpg vs hp using ggplot2
ggplot(mtcars, aes(x = hp, y = mpg)) +
    geom_point() +
    xlab("Horsepower") +
    ylab("Miles per Gallon") +
    ggtitle("Miles per Gallon vs Horsepower")
```